---
title: "Healthy ASD Control Samples Quality Control + Dimensionality Reduction Results (GTE)"
date: 2023-08-09
author: Samantha Yuen
output:
  html_document:
    toc: yes
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "./figs-gte/"
    )
  })
editor_options:
  chunk_output_type: inline
---

**Goal**: Optimizing formatting of figures for my thesis

## Load Libraries

```{r}
set.seed(100)

library(ggplot2)
library(SeuratObject)
library(Seurat)
library(ggbreak)
library(dplyr)

library(conflicted)
conflict_prefer("select", "dplyr") ## required in %>% dplyr

print(getwd())
size = 5

samplePalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#CC79A7", 
                   "#ff716e", "#999999", "#0072B2", "#194c76", "#D55E00", 
                   "#3a4f41", "#6699cc", "#713e5a")
    # "SRR9262920" "SRR9262922" "SRR9262923" "SRR9262932" "SRR9262937" 
    # "SRR9262938" "SRR9262946" "SRR9262956" "SRR9264382" "SRR9264383" 
    # "SRR9264385" "SRR9264388" "SRR9264389"


celltypePalette <- c("#E69F00", "#ff716e", "#CC79A7", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#A5B6AD")
celltypes <- c(
  "Astrocyte",
  "OPC",
  "Diff. Oligodendrocytes",
  "Microglia",
  "Proliferating",
  "Endothelial",
  "T-cell",
  "Intermediate")
```

## Load Data

```{r Load-GE}
data <- readRDS("D:\\backup files\\getegbm\\gete-gbm\\results\\GBMGSCTE3_Healthy\\2023-03-20_gbmgsc_healthy_raw\\gte_healthy_raw.rds") 
print(head(DefaultAssay(data)))
DefaultAssay(data) <- "RNA"

colnames(data@meta.data)

sampleIDs <- sort(unique(data$sampleID))
print(sampleIDs)

# Factor sample names
data$sampleID <- factor(data$sampleID, levels = sampleIDs)
```

## Quality Control figures

Compute Cell counts per sample before filtering

```{r}
df <- data@meta.data %>% select(sampleID) %>% count(sampleID) 
write.csv(df, file="healthy_gte_samplecounts.csv", row.names = F)
```

[ggbreak](https://www.frontiersin.org/articles/10.3389/fgene.2021.774846/full) was used to split bars in a bar graph.

```{r}
# Cell Count (Bar)
p <- data@meta.data %>%
    ggplot(aes(x=sampleID, fill=sampleID)) + 
    labs(fill="Samples") + labs(x = "", y = "Unfiltered Cell Count") +
    scale_fill_manual(values = samplePalette) +
    # scale_x_discrete(labels = sampleID) +
    # scale_fill_discrete(labels = sampleID) +
    geom_bar() +
    theme_classic() + scale_y_break(c(30000, 70000), scales = 0.1, ticklabels = c(70000)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.text = element_text(size=10), legend.position = "none")
p
ggsave("fig_gte_cellcount-bar.tiff", plot = p, units="in", width=size*1, height=size*0.6, dpi=300, compression = 'lzw')

## MitoRatio x UMI x Genes
p <- data@meta.data %>% 
    ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=mitoRatio)) + 
    geom_point() + 
    scale_colour_gradient(low = "gray90", high = "black") +
    stat_smooth(method=lm) +
    scale_x_log10() + 
    scale_y_log10() + 
    labs(x = "log10 UMI Detected per Cell", y = "log10 Genes Detected per Cell") +
    theme_classic() +
    geom_vline(xintercept = 1000, linetype = "dashed", colour = "red") +
    geom_hline(yintercept = 300, linetype = "dashed", colour = "red") +
    facet_wrap(~sampleID) + 
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=11, hjust=0.5), 
          strip.background = element_blank(),
          panel.border = element_rect(fill = NA, colour = "black"))
p
ggsave("fig_healthy_gte_umigenemitoratio_intercepts.tiff", plot = p, units="in", width=size*1.5, height=size*1.5, dpi=300, compression = 'lzw')

## MitoRatio
p <- data@meta.data %>% 
  	ggplot(aes(color=sampleID, x=mitoRatio, fill=sampleID)) + 
  	geom_density(alpha = 0.2) + 
    scale_color_manual(values = samplePalette, name = "Sample") +
    scale_fill_manual(values = samplePalette, name = "Sample") +
    labs(x = "Mitochondrial Expression (%)", y = "Cell Density") +
  	theme_classic() + scale_x_break(c(0.25, 0.6), scales = 0.1, ticklabels = c(0.60)) +
    theme(axis.title.x = element_text(size=11, hjust=0.35)) +
  	geom_vline(xintercept = 0.2, linetype = "dashed", colour = "red")
p
ggsave("fig_healthy_gte_mitoratio_density_legend.tiff", plot = p, units="in", width=size*1.3, height=size*0.8, dpi=300, compression = 'lzw')
ggsave("fig_healthy_gte_mitoratio_density.tiff", plot = p, units="in", width=size*1.1, height=size*0.6, dpi=300, compression = 'lzw')

## Genes per UMI (Density plot)
p <- data@meta.data %>% 
  	ggplot(aes(color=sampleID, x=log10GenesPerUMI, fill=sampleID)) + 
  	geom_density(alpha = 0.2) + 
    scale_color_manual(values = samplePalette) +
    scale_fill_manual(values = samplePalette) +
  	theme_classic() +
    labs(x = "log10 Genes Detected per Cell", y = "Cell Density") +
  	geom_vline(xintercept = 0.8, linetype = "dashed", colour = "red")
p
ggsave("fig_healthy_gte_log10GenesUMI_density.tiff", plot = p, units="in", width=size*1.1, height=size*0.6, dpi=300, compression = 'lzw')

```

Remove object from memory

```{r}
rm(data)
gc()
```

## QC Figures post QC, Normalization, and Clustering

Load Data

```{r Load-GE}
#load 
data <- readRDS("D:\\backup files\\getegbm\\gete-gbm\\results\\GBMGSCTE3_Healthy\\...\\ge_healthy_qc.rds") 
print(head(DefaultAssay(data)))
DefaultAssay(data) <- "RNA"

colnames(data@meta.data)

sampleIDs <- sort(unique(data$sampleID))
print(sampleIDs)

# Factor sample names
data$sampleID <- factor(data$sampleID, levels = sampleIDs)
```

### Compute Cell counts per sample after filtering

```{r}
df <- data@meta.data %>% select(sampleID) %>% count(sampleID) 
write.csv(df, file="healthy_gte_samplecounts.csv", row.names = F)
```

### PCA figures

```{r}
# PCA Elbow Plot
p <- ElbowPlot(data)
ggsave("fig_gte_elbowplot.tiff", plot = p, units="in", width=size*0.7, height=size*0.7, dpi=300, compression = 'lzw')

# Mito Ratio x Te Ratio

# Te Barplot

# Cell Count to csv
DefaultAssay(data) <- "RNA"
df <- data@meta.data %>% count(sampleID) 
df$genes.cell <- (dim(data)[1])/df$n
write.csv(df, file="healthy_gte_cellcounts.csv", row.names = F)

df <- data@meta.data %>% count(integrated_snn_res.0.3) 
write.csv(df, file="healthy_gte_cellcountscluster0.3.csv", row.names = F)

df <- data@meta.data %>% count(integrated_snn_res.0.4) 
write.csv(df, file="healthy_gte_cellcountscluster0.4.csv", row.names = F)

df <- data@meta.data %>% select(integrated_snn_res.0.3, gbm_subtype) %>% count(integrated_snn_res.0.3, gbm_subtype) 
write.csv(df, file="healthy_gte_cellcountssubtypes.csv", row.names = F)
```

## Cells per Cluster per Sample

```{r}
# PC20r03
DefaultAssay(data) <- "integrated"
df <- data@meta.data %>%
                group_by(sampleID) %>%
                count(integrated_snn_res.0.3)

df <- df %>% 
            group_by(integrated_snn_res.0.3) %>% 
            mutate(freq = (n / sum(n))*100) %>%
            mutate(rounded = round(freq, 2))
head(df)

p <- df %>% ggplot(aes(x=integrated_snn_res.0.3, y=rounded, fill=sampleID)) + 
        scale_fill_manual(values = samplePalette) +
        geom_bar( stat="identity", width = 0.9) + theme_classic() + 
        labs(y = "Cells per Cluster per Sample (%)", x = "hg38 + TE Mapped Cluster ID", fill= "Samples") +
        theme(axis.title = element_text(size=10), legend.position = "none")
ggsave("fig_gte_cellsxClusterxSamples_PC20r03.tiff", plot = p, units="in", width=size*0.6, height=size*0.6, dpi=300, compression = 'lzw')

# PC20r04
df <- data@meta.data %>%
                group_by(sampleID) %>%
                count(integrated_snn_res.0.4)

df <- df %>% 
            group_by(integrated_snn_res.0.4) %>% 
            mutate(freq = (n / sum(n))*100) %>%
            mutate(rounded = round(freq, 2))
head(df)


p <- df %>% ggplot(aes(x=integrated_snn_res.0.4, y=rounded, fill=sampleID)) + 
        scale_fill_manual(values = samplePalette) +
        geom_bar( stat="identity", width = 0.9) + theme_classic() + 
        labs(y = "Cells per Cluster per Sample (%)", x = "hg38 Mapped Cluster ID", fill= "Samples") +
        theme(axis.title = element_text(size=10), legend.position = "none")
ggsave("fig_gte_cellsxClusterxSamples_PC20r04.tiff", plot = p, units="in", width=size*0.6, height=size*0.6, dpi=300, compression = 'lzw')

# Celltype Ratio per Sample
```

## UMAPs

```{r}
# Clusters (resolution 0.3)
Idents(data) <- "integrated_snn_res.0.3"
DimPlot(data, label=TRUE) 
p <-DimPlot(data, label=FALSE) 
ggsave("fig_gte_PC20r03_UMAP.tiff", plot = p, units="in", width=size*1.1, height=size, dpi=300, compression = 'lzw')

# Clusters (resolution 0.4)
Idents(data) <- "integrated_snn_res.0.4"
head(Idents(data))
DimPlot(data, label=TRUE) 
p <- DimPlot(data, label=FALSE) 
ggsave("fig_gte_PC20r04_UMAP.tiff", plot = p, units="in", width=size*1.1, height=size*1, dpi=300, compression = 'lzw')

# Samples
Idents(data) <- "sampleID"
p <- DimPlot(data, label=FALSE, cols = samplePalette) 
ggsave("fig_gte_UMAP_samplesCombined.tiff", plot = p, units="in", width=size*1.2, height=size*1, dpi=300, compression = 'lzw')

# Cell Types
  # Move celltype labels to meta files
celltypeMeta <- read.csv("D:\\backup files\\getegbm\\gete-gbm\\results\\GBM-GSC neuroblastoma TE\\2022-11-07_celltypes_gbmscte\\gbmscte_meta_celltypes.csv", row.names = 1)
data@meta.data$int03_celltypes <- factor(celltypeMeta$int03_celltypes, levels = celltypes)
Idents(data) <- "int03_celltypes"
p <- DimPlot(data, label=FALSE, cols = celltypePalette) 
ggsave("fig_gte_UMAP_celltypes.tiff", plot = p, units="in", width=size*1.3, height=size*1, dpi=300, compression = 'lzw')

df <- data@meta.data %>% select(integrated_snn_res.0.3, int03_celltypes) %>% count(integrated_snn_res.0.3, int03_celltypes) 
write.csv(df, file="gte_celltypecounts.csv", row.names = F)

data@meta.data$int04_celltypes <- factor(celltypeMeta$int04_celltypes, levels = celltypes)
df <- data@meta.data %>% select(integrated_snn_res.0.4, ) %>% count(integrated_snn_res.0.3, int03_celltypes) 
write.csv(df, file="gte_celltypecounts.csv", row.names = F)

# GBM subtypes
data$gbm_subtype <- factor(data@meta.data$gbm_subtype, levels = gbmsubtypes, labels = gbmsubtypes_labels)
Idents(data) <- "gbm_subtype"
# data <- SetIdent(data, value = data@meta.data$gbm_subtype) 
p <- DimPlot(data, label=FALSE, cols = gbmPaletteGTE) 
ggsave("fig_gte_UMAP_gbmsubtypes.tiff", plot = p, units="in", width=size*1.1, height=size*1, dpi=300, compression = 'lzw')
```

## Dot Plots
